The Calculus of Linear Constructions

### 2021-12-02 writing/draft5/draft5.tex
writing

\subsection{Stateful Reasoning}
We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
\begin{definition}
  State axioms.
  \begin{itemize}
    \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U) \rightarrow A \rightarrow L$ \\
          The $\mapsto$ is a linear type constructor. It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion. Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
    \item $\textbf{new} : (A : U) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
          $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
    \item $\textbf{free} : (A : U) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \unit$ \\
          $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use. This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
    \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there. The type of $\get$'s output is the rather complicated $F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$. This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored. The proof $(l \mapsto_A m)$ is returned unchanged as well.
    \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
          $\set$ has the most complicated type of all the axioms presented here. Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term. This is the so called "strong update" operation. The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
  \end{itemize}
\end{definition}
Due to the fact that $(l \mapsto_A m)$ is a linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms. This is the same non-sharing principle that empowers Separation Logic. Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.

>>>>

\subsection{Stateful Reasoning}
We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
\begin{definition}
  State axioms.
  \begin{itemize}
    \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U) \rightarrow A \rightarrow L$ \\
          The $\mapsto$ is a linear type constructor. It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion. Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
    \item $\textbf{new} : (A : U) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
          $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
    \item $\textbf{free} : (A : U) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \unit$ \\
          $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use. This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
    \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there. The type of $\get$'s output is the rather complicated $F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$. This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored. The proof $(l \mapsto_A m)$ is returned unchanged as well.
    \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
          $\set$ has the most complicated type of all the axioms presented here. Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term. This is the so called "strong update" operation. The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
  \end{itemize}
\end{definition}
Due to the fact that $(l \mapsto_A m)$ is an abstract linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms. This is the same non-sharing principle that empowers Separation Logic. Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
